---
title: "Rshiny"
output: html_document
---

```{r}
library(shiny)
library(bslib)
library(dplyr)
library(ggplot2)
library(ggExtra)
library(sf)
library(leaflet)
library(openxlsx)
library(here)
library(lubridate)
library(plotly)
library(patchwork)
```

```{r}
#read in the dataset of imidacloprid

#try reading in all of the datasets at once, as a excel spreadsheet
#this reads in all of the sheets at once!
file_in <- here("data", "Pharmaceuticals infomation_V3.xlsx")

# Get the name of the sheets.
f_sheets <- getSheetNames(file_in)
 
# Read in the ibuprofen data
imidacloprid <- read.xlsx(file_in, sheet = "Imidacloprid_LCMS")

#store year as an integer
imidacloprid$year <- as.integer(imidacloprid$year)

#store datetime as datetime
imidacloprid$Sample_datetime <- ymd_hms(imidacloprid$Sample_datetime)

```

```{r}
#read in the rainfall dataset - this was created from the 'extracting site rainfall data' script
rainfall <- read.csv(here("data", "daily_rainfall_1km", "rainfall.csv"))

#save the date as a date format
rainfall$date <- as.Date(rainfall$date)
```


```{r}
#i want the columns concentration, year, longitude and latitude. also convert the coordinates to the sf values so I can plot them

imidacloprid_sf <- imidacloprid %>%
  select(Concentration, year, Longitude, Latitude, Sample_Site_ID, Sample_datetime) %>% #select the columns that I want
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) #%>% #converts coordinate column into a geometry column
  #mutate(log_Concentration = log(Concentration))

#because I am plotting by sample site ID, i need to remove sites where the site ID has been redacted. 
#otherwise it gets really confusing with the same sample site ID have multiple coordinates
imidacloprid_sf <- imidacloprid_sf %>%
  filter(Sample_Site_ID != "REDACTED")

#for use in another script, i want a list of unique sites and their coordinates
#unique_sites <- imidacloprid_sf %>%
 # dplyr::distinct(Sample_Site_ID, geometry)

#save as st file in the data folder
#st_write(unique_sites,here("data", "imidacloprid sites", "unique_sites.gpkg"),delete_dsn = TRUE)

```

############################################
legend where the size of the points shows the number of samples at each site
```{r}
#to adjust the size of the point by how well sampled it is, need to count the number of obersvations at each site.
site_counts <- imidacloprid_sf %>%
  st_drop_geometry() %>%
  count(Sample_Site_ID, name = "n_obs") %>% #count the number of observations at each site
  left_join( #join these counts onto the imidacloprid_sf dataset
    imidacloprid_sf %>%
      group_by(Sample_Site_ID) %>%
      summarise(geometry = st_union(geometry), .groups = "drop"),
    by = "Sample_Site_ID"
  ) %>%
  st_as_sf() #convert into a geometry column


#define the bins for the legend, and the radii of the points corresponding to the bins
legend_df <- data.frame(
  obs_bin = c("<10", "10–50", "50–100", "100–200",
              "200–300", "300–400", "400+"),
  radius = c(4, 6, 12, 14, 16, 18, 20)
)

#add these bins onto the site counts dataframe
site_counts <- site_counts %>%
  mutate(
    obs_bin = cut(
      n_obs,
      breaks = c(0, 10, 50, 100, 200, 300, 400, Inf),
      labels = legend_df$obs_bin,
      right = FALSE
    )
  ) %>%
  left_join(legend_df, by = "obs_bin")

```

```{r}
#build a html legend
legend_items <- paste0(
  sprintf(
    '<div style="display:flex;align-items:center;margin-bottom:6px;">
       <div style="
         width:%dpx;
         height:%dpx;
         border-radius:50%%;
         background:blue;
         opacity:0.7;
         margin-right:8px;
         flex-shrink:0;">
       </div>
       <span>%s</span>
     </div>',
    legend_df$radius,
    legend_df$radius,
    legend_df$obs_bin
  ),
  collapse = ""
)

legend_html <- paste0(
  '<div style="background:white;padding:8px 10px;border-radius:4px;">
     <strong>Number of samples</strong><br><br>',
  legend_items,
  '</div>'
)
```



```{r}  

ui <- fluidPage(

  titlePanel("Imidacloprid Concentrations"),
#controls the size of the left-hand column, which shows the map
  fluidRow(
    column(
      width = 7,
      leafletOutput("map", height = 500)
    ),
#controls the size of the right-hand column, which shows the timeseries plots
    column(
      width = 5,
      plotlyOutput("timeseries", height = 700)
    )
  )
)

server <- function(input, output, session) {

  # store clicked site
  selected_site <- reactiveVal(NULL)

  # initialise map ONCE
  output$map <- renderLeaflet({
    leaflet(imidacloprid_sf) %>%
      addProviderTiles(providers$CartoDB.Positron)
  })

  # handle clicks
  observeEvent(input$map_marker_click, {
    selected_site(input$map_marker_click$id)
  })

  # draw ALL points 
  observe({
    leafletProxy("map", data = site_counts) %>%
      clearMarkers() %>%
      clearControls() %>%
      addCircleMarkers(
        radius = ~radius,
        stroke = FALSE,
        fillColor = "blue",
        fillOpacity = 0.7,
        layerId = ~Sample_Site_ID,
        popup = ~paste("<b>Site:</b>", Sample_Site_ID, "<br>",
                       "<b>Samples:</b>", n_obs)
        )%>%
      addControl(
    html = HTML(legend_html),
    position = "bottomright"
  )
  })

  # make a concentration plot for the selected site
  site_data <- reactive({

    req(selected_site())

    imidacloprid_sf %>%
      st_drop_geometry() %>%
      filter(Sample_Site_ID == selected_site()) %>%
      arrange(Sample_datetime)

  })

  #make a rainfall plot for the selected site
  rain_data <- reactive({

  req(selected_site())

  dat <- rainfall %>%
    filter(Sample_Site_ID == selected_site()) %>%
    arrange(date)


})
  
 output$timeseries <- renderPlotly({

  req(selected_site())

  p_rain <- plot_ly(
    data = rain_data(),
    x = ~date,
    y = ~rainfall,
    type = "bar",
    marker = list(color = "steelblue"),
    hovertemplate =
      paste(
        "<b>Date:</b> %{x}<br>",
        "<b>Rainfall:</b> %{y:.1f} mm",
        "<extra></extra>"
      )
  ) %>%
    layout(
      yaxis = list(title = "Rainfall (mm)")
    )

  p_conc <- plot_ly(
    data = site_data(),
    x = ~Sample_datetime,
    y = ~Concentration,
    type = "scatter",
    mode = "lines+markers",
    line = list(color = "black"),
    marker = list(size = 5),
    hovertemplate =
      paste(
        "<b>Date:</b> %{x}<br>",
        "<b>Concentration:</b> %{y:.3f} µg/L",
        "<extra></extra>"
      )
  ) %>%
    layout(
      title = paste("Site:", selected_site()),
      xaxis = list(title = "Date"),
      yaxis = list(
        title = "Imidacloprid (µg/L)",
        range = c(0.0002, 0.25)
      )
    )

  subplot(
    p_rain,
    p_conc,
    nrows = 2,
    shareX = TRUE,
    heights = c(0.5, 0.5),
    titleY = TRUE
  ) %>%
    layout(
      showlegend = FALSE,
      hovermode = "x unified"
    )

})


}

shinyApp(ui, server)
```

###########################################################################
legend by colour AND size
the size of the point shows the median number of samples per year
and the colour of the point shows for how many years we have data for
```{r}
#calculate the median number of samples per year at each site, and also the number of years that each site has been sampled.
site_summaries <- imidacloprid_sf %>%
  group_by(Sample_Site_ID, year) %>%
  summarise(
    annual_samples = n(),
    .groups = "drop"
  ) %>%
  group_by(Sample_Site_ID) %>%
  summarise(
    median_annual_samples = median(annual_samples),
    years_sampled = n_distinct(year),
    .groups = "drop"
  )



#define the size of bins for the size of the point (median number of samples per year)
size_legend_df <- data.frame(
  sample_bin = c("1", "2–5", "6–10", "11–15", "16–20", "21–25", "26–29"),
  radius = c(4, 6, 8, 10, 12, 16, 20)
)

#assign the bins to the site summaries
site_summaries <- site_summaries %>%
  mutate(
    sample_bin = cut(
      median_annual_samples,
      breaks = c(0, 2, 6, 11, 16, 21, 26, Inf),
      labels = size_legend_df$sample_bin,
      right = FALSE
    )
  ) %>%
  left_join(size_legend_df, by = "sample_bin")




#define colour bins
year_breaks <- c(0, 2, 4, 6, 8, 10, Inf)
year_labels <- c("1", "2–3", "4–5", "6–7", "8–9", "10–11")

site_summaries <- site_summaries %>%
  mutate(
    year_bin = cut(
      years_sampled,
      breaks = year_breaks,
      labels = year_labels,
      right = FALSE
    )
  )

#create a palette
pal <- colorFactor(
  palette = "YlOrRd",
  levels = year_labels
)

```


```{r}
#size legend
legend_items_size <- paste0(
  sprintf(
    '<div style="display:flex;align-items:center;margin-bottom:6px;">
       <div style="
         width:%dpx;
         height:%dpx;
         border-radius:50%%;
         background:grey;
         opacity:0.7;
         margin-right:8px;">
       </div>
       <span>%s</span>
     </div>',
    size_legend_df$radius,
    size_legend_df$radius,
    size_legend_df$sample_bin
  ),
  collapse = ""
)

legend_html_size <- paste0(
  '<div style="background:white;padding:8px;border-radius:4px;">
  <strong>Median samples per year</strong><br><br>',
  legend_items_size,
  '</div>'
)


#colour legend
legend_items_colour <- paste0(
  sprintf(
    '<div style="display:flex;align-items:center;margin-bottom:4px;">
       <div style="
         width:18px;
         height:18px;
         background:%s;
         margin-right:8px;">
       </div>
       <span>%s</span>
     </div>',
    pal(year_labels),
    year_labels
  ),
  collapse = ""
)

legend_html_colour <- paste0(
  '<div style="background:white;padding:8px;border-radius:4px;">
  <strong>Number of years sampled</strong><br><br>',
  legend_items_colour,
  '</div>'
)
```


this app has the new legend and just concentration, no rainfall

```{r} 

#thought: could I colour the points based on whether imidacloprid concentration is above or below a safe threshold?
#or maybe add a line to the timseries graph, to show this threshold?
ui <- fluidPage(

  titlePanel("Imidacloprid Concentrations"),

  fluidRow(
    column(
      width = 7,
      leafletOutput("map", height = 500)
    ),

    column(
      width = 5,
      plotlyOutput("timeseries", height = 500)
    )
  )
)

server <- function(input, output, session) {

  # store clicked site
  selected_site <- reactiveVal(NULL)

  # initialise map ONCE
  output$map <- renderLeaflet({
    leaflet(imidacloprid_sf) %>%
      addProviderTiles(providers$CartoDB.Positron)
  })

  # handle clicks
  observeEvent(input$map_marker_click, {
    selected_site(input$map_marker_click$id)
  })

  # draw ALL points 
  observe({
    
    leafletProxy("map", data = site_summaries) %>%
  clearMarkers() %>%
  clearControls() %>%
  addCircleMarkers(
    radius = ~radius,
    fillColor = ~pal(year_bin),
    fillOpacity = 0.8,
    stroke = FALSE,
    layerId = ~Sample_Site_ID,
    popup = ~paste(
      "<b>Site:</b>", Sample_Site_ID, "<br>",
      "<b>Median annual samples:</b>", median_annual_samples, "<br>",
      "<b>Years sampled:</b>", years_sampled
    )
  ) %>%
  addControl(HTML(legend_html_size), position = "bottomright") %>%
  addControl(HTML(legend_html_colour), position = "topright")
    
    
    
    
    
  })

  # time series data for selected site
  site_data <- reactive({

    req(selected_site())

    imidacloprid_sf %>%
      st_drop_geometry() %>%
      filter(Sample_Site_ID == selected_site()) %>%
      arrange(Sample_datetime)

  })

  

  # plot time series
  output$timeseries <- renderPlotly({
    

    req(selected_site())

    p <- ggplot(site_data(), aes(x = Sample_datetime,
                           y = Concentration
                           )) +
   geom_line() +
    geom_point(size = 1) +

  scale_y_continuous(
    limits = c(0.0002, 0.25),
    breaks = c(0, 0.05, 0.1, 0.15, 0.20, 0.25)) +
      labs(
        title = paste("Site:", selected_site()),
        x = "date",
        y = "Imidacloprid concentration (ug/l)"
      ) +
      theme_bw()+
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
    


  })

}

shinyApp(ui, server)
```

trying out the new legend along with the rainfall plotted
trying out with two check boxes
```{r}  

ui <- fluidPage(

  titlePanel("Imidacloprid Concentrations"),
#this is the size of the first column, which shows the map
  fluidRow(
    column(
      width = 7,
      leafletOutput("map", height = 500)
    ),
#this is the size of the second column, which shows the concentration and rainfall timeseries
    column(
      width = 5,
      plotlyOutput("timeseries", height = 500),
     checkboxInput(        #add a checkbox for MAc-EQS
  "show_mac_eqs",
  HTML(
    "Show MAC-EQS (0.057 µg/L)<br>
    <span style='font-size:90%; color:#666;'>
      Short-term exposure threshold
    </span>"
  ),
  value = FALSE
),

checkboxInput(
  "show_aa_eqs",       #add a checkbox for AA-EQS
  HTML(
    "Show AA-EQS (0.0068 µg/L)<br>
    <span style='font-size:90%; color:#666;'>
      Long-term exposure threshold
    </span>"
  ),
  value = FALSE
)
      
    )
  )
)

server <- function(input, output, session) {

  # store clicked site
  selected_site <- reactiveVal(NULL)

  # initialise map ONCE
  output$map <- renderLeaflet({
    leaflet(imidacloprid_sf) %>%
      addProviderTiles(providers$CartoDB.Positron)
  })

  # handle clicks
  observeEvent(input$map_marker_click, {
    selected_site(input$map_marker_click$id)
  })

  # draw ALL points 
  observe({
leafletProxy("map", data = site_summaries) %>%
  clearMarkers() %>%
  clearControls() %>%
  addCircleMarkers(
    radius = ~radius,
    fillColor = ~pal(year_bin),
    fillOpacity = 0.8,
    stroke = FALSE,
    layerId = ~Sample_Site_ID,
    popup = ~paste(
      "<b>Site:</b>", Sample_Site_ID, "<br>",
      "<b>Median samples per year:</b>", median_annual_samples, "<br>",
      "<b>Number of years sampled:</b>", years_sampled
    )
  ) %>%
  addControl(HTML(legend_html_size), position = "bottomright") %>%
  addControl(HTML(legend_html_colour), position = "topright")
  })

  # make a concentration plot for the selected site
  site_data <- reactive({

    req(selected_site())

    imidacloprid_sf %>%
      st_drop_geometry() %>%
      filter(Sample_Site_ID == selected_site()) %>%
      arrange(Sample_datetime)

  })

  #make a rainfall plot for the selected site
  rain_data <- reactive({

  req(selected_site())

  dat <- rainfall %>%
    filter(Sample_Site_ID == selected_site()) %>%
    arrange(date)


})
  
 output$timeseries <- renderPlotly({

  req(selected_site())

   #make the rainfall plot for the selected site
  p_rain <- plot_ly(
    data = rain_data(),
    x = ~date,
    y = ~rainfall,
    type = "bar",
    marker = list(color = "steelblue"),
    hovertemplate =
      paste(
        "<b>Date:</b> %{x}<br>",
        "<b>Rainfall:</b> %{y:.1f} mm",
        "<extra></extra>"
      )
  ) %>%
    layout(
      yaxis = list(title = "Rainfall (mm)")
    )
#make a concentration plot for the selected site
 p_conc <- plot_ly(
  data = site_data(),
  x = ~Sample_datetime,
  y = ~Concentration,
  type = "scatter",
  mode = "lines+markers",
  line = list(color = "black"),
  marker = list(size = 5, color = "black"),
  hovertemplate =
    paste(
      "<b>Date:</b> %{x}<br>",
      "<b>Concentration:</b> %{y:.3f} µg/L",
      "<extra></extra>"
    )
)

# Base layout
p_conc <- p_conc %>%
  layout(
    title = paste("Site:", selected_site()),
    xaxis = list(title = "Date"),
    yaxis = list(
      title = "Imidacloprid (µg/L)",
      range = c(0.0002, 0.25)
    )
  )

# Add MAC-EQS line if requested
shapes <- list()
annotations <- list()

# MAC-EQS
if (input$show_mac_eqs) {
  
#plots the line
  shapes <- append(shapes, list(
    list(
      type = "line",
      xref = "paper",
      x0 = 0,
      x1 = 1,
      y0 = 0.057,
      y1 = 0.057,
      line = list(
        color = "red",
        dash = "dash",
        width = 2
      )
    )
  ))
  
#adds an annotation to the line
  annotations <- append(annotations, list(
    list(
      x = 0.98,
      xref = "paper",
      y = 0.065,
      yref = "y",
      text = "MAC-EQS",
      showarrow = FALSE,
      xanchor = "right",
      font = list(color = "red")
    )
  ))
}

# AA-EQS
if (input$show_aa_eqs) {

  #plots the line, in green
  shapes <- append(shapes, list(
    list(
      type = "line",
      xref = "paper",
      x0 = 0,
      x1 = 1,
      y0 = 0.0068,
      y1 = 0.0068,
      line = list(
        color = "darkgreen",
        dash = "dash",
        width = 2
      )
    )
  ))

  #adds an annotation to the line
  annotations <- append(annotations, list(
    list(
      x = 0.98,
      xref = "paper",
      y = 0.012,
      yref = "y",
      text = "AA-EQS",
      showarrow = FALSE,
      xanchor = "right",
      font = list(color = "darkgreen")
    )
  ))
}

# Apply both lines if both boxes are ticked
p_conc <- p_conc %>%
  layout(
    shapes = shapes,
    annotations = annotations
  )

  
#adds the vertical line to scroll along the x-axis
  subplot(
    p_rain,
    p_conc,
    nrows = 2,
    shareX = TRUE,
    heights = c(0.5, 0.5),
    titleY = TRUE
  ) %>%
    layout(
      showlegend = FALSE,
      hovermode = "x unified"
    )

})


}

#display the shiny app
shinyApp(ui, server)
```







